---
title: "association analysis"
author: "Quincy"
date: "9/9/2021"
output:
  pdf_document: default
  html_document: default
---
# Association Analysis

## Context
Create association rules that will allow us to identify relationships between variables in the dataset. We have been provided with a dataset that comprises of  groups of items that will be associated with others. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# We first we install the required arules library 
#install.packages("arules")
```


```{r}
# Loading the arules library
library(arules)
```


# Load data

```{r}
path <- "http://bit.ly/SupermarketDatasetII"

super <- read.transactions(path, sep = ",")
super
```

```{r}
# Verifying the object's class
# This should show us transactions as the type of data that we will need

class(super)
```


```{r}
# Previewing our first 5 transactions
inspect(super[1:5])
```




```{r}
# Generating a summary of the supermarket dataset
# This would give us some information such as the most purchased items, 
# distribution of the item sets (no. of items purchased in each transaction), etc.

summary(super)
```

We can see that mineral water leads in sales followed by eggs,sphagetti,french fries and chocolate.



```{r}
# Exploring the frequency of some articles 
# i.e. transacations ranging from 8 to 10 and performing 
# some operation in percentage terms of the total transactions 

itemFrequency(super[, 8:10],type = "absolute")
round(itemFrequency(super[, 8:10],type = "relative")*100,2)
```

```{r}
# Displaying top 10 most common items in the transactions dataset 
# and the items whose relative importance is at least 10%
# 
par(mfrow = c(1, 2))

# plot the frequency of items
itemFrequencyPlot(super, topN = 10,col="darkgreen")
itemFrequencyPlot(super, support = 0.1,col="darkred")
```

```{r}
# Building a model based on association rules 
# using the apriori function 
# We use Min Support as 0.001 and confidence as 0.8

rules <- apriori (super, parameter = list(supp = 0.001, conf = 0.8))
rules
```

Since we built the model using 0.001 Min support and confidence as 0.8 we obtained 74 rules.

```{r}
# we will see what happens if we increase the support or lower the confidence level

# Building a apriori model with Min Support as 0.002 and confidence as 0.8.
rules2 <- apriori (super,parameter = list(supp = 0.002, conf = 0.8)) 
rules2

```
This gives us 2 rules which is too little.

```{r}
# Building apriori model with Min Support as 0.002 and confidence as 0.6.
rules3 <- apriori (super, parameter = list(supp = 0.001, conf = 0.6)) 
rules3
```
This gives us 545 rules which is okay but we will use rules.

```{r}
# More statistical information such as support, lift and confidence is also provided.

summary(rules)
```

Most rules have 3 and 4 items though some rules do have upto 5 and 6.

```{r}
# Observing rules built in our model i.e. first 5 model rules

inspect(rules[1:5])
```

If someone buys frozen smoothie and spinach they are 89% likely to buy mineral water.

```{r}
# Ordering the rules by a criteria such as the level of confidence
# then looking at the first five rules.

rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:5])
```

The first four rules have a confidence of 100%.

```{r}
# If we're interested in making a promotion relating to the sale of milk, 
# we could create a subset of rules concerning these products 
# This would tell us the items that the customers bought before purchasing mineral
milk <- subset(rules, subset = rhs %pin% "milk")
# Then order by confidence
milk <-sort(milk, by="count", decreasing=TRUE)
inspect(milk[1:5])
```
 
- meatballs,whole wheat pasta had been in 10 baskets.

- black tea,frozen smoothie had been in 9 baskets.

- cake,meatballs,mineral water had been in 8 baskets.

- escalope,hot dogs,mineral water had been in 8 baskets.

- burgers,ground beef,olive oil had been in 8 baskets.
